### 2018-03-11 hw1/hw1_backup.tex

\documentclass{article}

\usepackage{geometry}
\geometry{margin=1in}

\usepackage{pgfplots}

\title{STAT 672: Homework \#1}
\author{Tom Wallace}

\setlength{\parindent}{0cm}

\begin{document}

\maketitle

\section*{Problem 1}

\subsection*{A}

See attached code \texttt{hw1\_A.py} for an implementation of rejection sampling
from the unit cube.

\smallskip

The acceptance probability is the volume of an $n$-dimensional ball divided by
the volume of an $n$-dimensional cube:

\begin{equation}
P_{X \sim B^d_\infty}(||X||_2 \leq 1) =
\frac{\mathrm{Vol}(B^d_2)}{\mathrm{Vol}(B^2_\infty)}
= \frac{\frac{\pi ^ {d/2}}{\Gamma(\frac{d}{2} + 1)}}{2^d}
\end{equation}

We know that $\Gamma (n+1) = n!$ Thus, $\Gamma (\frac{d}{2} + 1) = (\frac{d}{2})!$

\smallskip

Stirling's approximation for factorials is useful here.

\begin{equation}
n! \approx \sqrt{2 \pi n}n^ne^{-n}
\end{equation}

Let us consider the case of $2d$. In this case:

\begin{equation}
\mathrm{Vol}(B^{2d}_2) = \frac{\pi ^ d}{\Gamma(d + 1)} = \frac{\pi^d}{d!}
\approx \frac{\pi^d}{\sqrt{2\pi d}d^de^{-d}}
\end{equation}

Rearranging terms, we have:

\begin{equation}
= \frac{1}{\sqrt{2\pi d}} \left( \frac{\pi e}{d} \right)^d
\end{equation}

Referring to $ \left( \frac{\pi e}{d} \right)^d$, the denominator grows faster with $d$ than does the
numerator. We
can show this more formally by computing the derivative of the numerator and of the
denominator, respectively:

\begin{equation}
\frac{d}{dx} ((\pi e)^x) = (\pi e)^x (1 + \ln \pi)
\end{equation}

\begin{equation}
\frac{d}{dx}(x^x) = x^x (1 + \ln x)
\end{equation}

As $x$ grows larger, (6) becomes increasingly larger than (5). As a consequence, if we extend $d$ infinitely, 
the limit of the ratio is 0.

\begin{equation}
\lim_{d \to \infty} \left( \frac{\pi e}{d} \right)^d = 0
\end{equation}

Which implies:
\begin{equation}
\lim_{d \to \infty} \frac{1}{\sqrt{2\pi d}} \left( \frac{\pi e}{d} \right)^d = 0
\end{equation}

In other words, as $d \to \infty$, more and more of the volume of the unit ball
is contained near its surface. This phenomenon is called \textit{concentration
of measure}.

\smallskip

Returning to (1), we compute the the volume of the unit cube as $d \to \infty$.

\begin{equation}
\lim_{d \to \infty} 2^{2d} = \infty
\end{equation}

So, the ratio of the unit sphere to the unit cube is:

\begin{equation}
\lim_{d \to \infty} \frac{\mathrm{Vol}(B^d_2)}{\mathrm{Vol}(B^2_\infty)} =
\frac{0}{\infty} = 0
\end{equation}

Since this ratio also is the probability of accepting a sample in a rejection sampling
scheme, our expected runtime to generate a single random vector from $\mathrm{B}^d_2$ with
rejection sampling grows asymptotically with $d$. For example, for $d=10000$, our
program would run for a very, very, \textit{very} long time.

\subsection*{B}
We can generate $\Theta$ as $\Theta = Z / ||Z||_2$, where $Z \sim N(0, I_d)$ due
to the Box-Muller (B-M) transformation.

\section*{Problem 2}

\subsection*{A}

Below are the results of my simulation, the code of which is attached as
\texttt{hw1\_B.py}. As $d$ grows larger, the frequency of the event $\{||X_+ -
Z||_2 \geq ||X_{-} - Z||_2\}$ increases. This is somewhat counter-intuitive.
Because $Z$ is drawn from the same distribution as $X_+$, we would expect their
difference $X_+ - Z$ to result in something close to a zero-vector, which should
be smaller than the difference between $X_-$ and $Z$ (since they have different
means and hence should have a larger difference). Yet, as dimensionality grows,
this is not the case: the event occurs with increasing frequency. This has two
explanations. One, because the mean is $\pm 5/\sqrt{d}$, as $d$ increases, the
means increasingly converge towards 0. Two, $X_+$ and $Z$ have larger variance than
$X_-$, and this variance does not decrease with $d$. Thus, as $d$ increases, the
larger variance of $X_+$ and $Z$ predominates over their increasingly similar
means, resulting in them more often being further apart (their difference has a
larger Euclidean norm) than $X_-$ and $Z$.

\begin{tikzpicture}
\begin{axis}[ylabel = Number of Events, 
	     xlabel = Dimensionality, 
	     xbar interval = 0.8,
\addplot[fill=blue] coordinates { (foo, 1) (bar, 2)};
\end{axis}
\end{tikzpicture}

\end{document}