### 2018-03-11 week1/hw1.tex

\documentclass{article}
\begin{document}

\section{Problem 1}

$f(x) = ||\mathbf{Ax-b}||^2_2$

\medskip

This is just the ordinary least squares OLS problem

$= (\mathbf{Ax-b)'(Ax-b)}$

$= (\mathbf{x'A'-b')(Ax-b)}$

$= \mathbf{x'A'Ax - x'A'b - b'Ax + b'b}$

$= \mathbf{x'A'Ax - 2x'A'b + b'b}$

\medskip

$\nabla f(x) = \mathbf{2A'Ax-2A'b}$

$\nabla^2 f(x) = \mathbf{2A'A}$

\medskip

Reminders:

$\mathbf{Av = \lambda v}$

$\mathbf{v}$ is an eigenvector if it satisfies this equation

$\lambda$ is the eigenvalue corresponding to $\mathbf{v}$

Eigenvectors are the vectors that linear transformation $\mathbf{A}$ merely elongates or shrinks (as opposed to
rotating)

Eigenvalue is the amount that they elongate or shrink

To find, solve $det(\mathbf{A-\lambda I})=0$ (characteristic equation)

Eigendecomposition: $\mathbf{A=Q \Lambda Q^{-1}}$

$\mathbf{Q}$ is a $n \times n$ matrix whose $i$'th column is eigenvector $q_i$ of $\mathbf{A}$

$\Lambda$ is a diagonal matrix for which $\Lambda_{ii}=\lambda_i$

If eigenvalues of Hessian are all positive, then $\mathbf{A}$ is positive semi-definite (multivariate equivalent of
convex)

If all negative, then $\mathbf{A}$ is negative-definite (multivariate equivalent of concave)

Matrix $\mathbf{A}$ is positive definite IFF $\mathbf{A}=\mathbf{R}'\mathbf{R}$, where $\mathbf{R}$ is a matrix with
independent columns (full rank)

\medskip

So, in our case, assuming $\mathbf{A}$ is of full rank, since it has a Hessian of $\mathbf{A}\mathbf{A}'$, it is positive
definite and so all eigenvalues are positive.

\medskip

The necessary and sufficient condition for $x^*$ to be a minimizer of $f(x)$ is that the gradient is 0 at $x^*$

\medskip

$f$ has a unique minimizer if $A$ is positive definite

\section{Problem 2}

By the law of total probability, $ f_{X}(x) = \sum_{k=1}^K f_{X|Y=k}(x|k) f_Y(k)$

\medskip

$P(Y=k|X=x) = \frac{P(X=x|Y=k)P(Y=k)}{P(X=x|Y=k)P(Y=k)+P(X=x|Y\neq k)P(Y\neq k)}$
$=\frac{f_{X|Y=k}(x)P(Y=k)}{f_X(x)}$
\end{document}