### 2018-04-01 hw4/hw4.tex
finished hw4

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{geometry}
\geometry{margin=1.1in}

\title{STAT 672: Homework 4}
\author{Tom Wallace}

\begin{document}

\maketitle

\section*{A}

See \texttt{hw4.py} for code.

It is recommended to scale data when using SVM because if we do not, we
essentially penalize each feature differently. To quote Hsu, Chang, and Lin
(2008), "the main advantage [of scaling] is to avoid attributes in greater
numeric ranges dominating those in smaller numeric ranges" \cite{hsu}.

This can be shown analytically. Consider a typical soft-margin SVM
classification problem.
$$
\min_{w_0, w} C \frac{1}{n} \sum_{i=1}^n \max(0, 1 - Y_i(w_0 +
\langle w, \Phi(X_i) \rangle)) + \frac{1}{2}||w||^2_2
$$

Let $\lambda = \frac{1}{2C}$ and assume for convenience with no loss of
generality that centering has occurred and so no intercepts are needed.

$$
\min_{w} \frac{1}{n} \sum_{i=1}^n \max(0, 1 - Y_i(
\langle w, \Phi(X_i) \rangle)) + \lambda||w||^2_2
$$

Consider a diagonal matrix $S$ whose non-zero entries consist of the
respective L2 norms of the columns of $\Phi$:

$$
S := \mathrm{diag}(||\Phi_{\bullet 1}||_2 \ldots ||\Phi_{\bullet D}||_2)
$$

Restate the SVM using $S$:

$$
\min_{w} \frac{1}{n} \sum_{i=1}^n \max(0, 1 - Y_i(
\langle wS, S^{-1}\Phi(X_i) \rangle)) + \lambda||w||^2_2
$$

Define $\theta = wS$:
$$
\min_{\theta} \frac{1}{n} \sum_{i=1}^n \max(0, 1 - Y_i(
\langle \theta, S^{-1}\Phi(X_i) \rangle)) + \lambda||\theta S^{-1}||^2_2
$$

We now can see that feature with a small numerical scale are penalized more
heavily (conversely, features with a large numerical scale will have greater
emphasis in the final classifier).

$$
= 
\min_{\theta} \frac{1}{n} \sum_{i=1}^n \max(0, 1 - Y_i(
\langle \theta, S^{-1}\Phi(X_i) \rangle)) + \lambda \sum_{j=1}^D
\frac{w_j^2}{||\Phi_{\bullet j}||^2_2}
$$

\section*{B}

See \texttt{hw4.py} and \texttt{output} for code and output.

\section*{C}

The number of support vectors decreases as C increases, as depicted in Table
1.

Explanation: C is the penalty on slack variables $\zeta$, which capture the
degree to which the margin constraint is violated. A small value of C
means a small penalty on margin violations, and will result in a relatively wide margin 
and many points laying on or inside of the margin (and hence a high number of support vectors). 
A large value of C means a
large penalty on margin violations, and so the separating hyperplane will have a
much tighter margin, meaning fewer
points that lie on or in the margin and hence a smaller number of support
vectors.

The above explanation also covers why run time for training increases with C,
eventually resulting in non-convergence.
When C is small, the separating
hyperplane is easy to find since margin violations are only
lightly punished, and so training run times are low. As C becomes larger, the
SVM tries more and more to find a separating hyperplane that doesn't result in
margin violations, and requires more calculation (run time) to find. 
Once C attains a sufficiently large value, the SVM may fail to converge because 
margin violations are unacceptable (due to the high value of C) but the data are 
not linearly separable and so some margin violation must be accepted, resulting
in a catch-22 that the optimization routine cannot overcome.

\begin{table}[h!]
	\caption{Influence of C on number of support vectors}
	\centering
	\begin{tabular}{|r r|}
		\hline
		\textbf{C} & \textbf{n(SV)} \\
		$10^{-1}$ & 1818 \\
		$10^{-0.5}$ & 1777 \\
		$10^{0}$ & 1443 \\
		$10^{0.5}$ & 1083 \\
		$10^{1}$ & 847 \\
		$10^{1.5}$ & 676 \\
		$10^{2}$ & 576 \\
		$10^{2.5}$ & 511 \\
		$10^{3}$ & 478 \\
		$10^{3.5}$ & 456 \\
		$10^{4}$ & 448 \\
		$10^{4.5}$ & 448 \\
		$10^{5}$ & 439 \\
		$10^{5.5}$ & 438 \\
		$10^{6}$ & 441 \\
		\hline
	\end{tabular}
\end{table}

\section*{D}

$w^*$ can be interpreted, but one must be cautious. A positive $w$
means the corresponding feature is associated with the positive class, while a
negative $w$ means the feature is associated with the negative class. The size
of $w$, within the context of a particular model, indicates the magnitude of
this association (e.g., if one $w$ is much larger in absolute value than other $w$, it is
strongly associated with a class, and probably is very useful for prediction). 
However, the caveat ``within a particular model'' is
crucial. Different regularization parameters will produce different coefficients
for identical datasets. We cannot make the sort of specific
interpetation that is possible with some parametric models, e.g. in linear
regression ``every one inch increase in height $X_1$ is associated with a $\beta_1$
increase in weight (lbs)."

Table 2 depicts the five largest and five smallest weights obtained with
$C=10^{4.5}$. We can interpret these weights. Emails that TALK IN ALL CAPS and
mention money (U.S. dollars) are especially likely to be categorized as spam.
Emails in which the sender mentions the name of someone that we presumably both
know (George) and refers to meeting
(presumably in-person) are indicative of a ``real'' relationship and hence 
are especially likely to be categorized as not-spam.

\begin{table}[h!]
	\caption{Largest and smallest weights}
	\centering
	\begin{tabular}{|l r | l r|}
		\hline
		\textbf{Feature} & \textbf{Weight} & \textbf{Feature} &
		\textbf{Weight} \\
		3d & 177.47 & george & -342.13 \\
		capital\_run\_length\_average & 144.75 & hp & -162.41 \\
		\$ & 45.77 & 415 & -161.59 \\
		857 & 33.01 & cs & -140.11 \\
		000 & 30.43 & meeting & -56.00 \\
		\hline
	\end{tabular}
\end{table}

\section*{E}

As shown in Table 3, relatively small values of C are associated with false
positive rates about 1\%. If we feel very strongly about staying below 1\%,
C=$10^{-0.5}$ has a FP rate of about 0.05\%. If we are willing to go slightly
above 1\%, C=$10^{0}$ has a FP rate of about 2\% and has over double the TP
rate.

\begin{table}[h!]
	\caption{True and false positive rates for selected C}
	\centering
	\begin{tabular}{|r r r|}
		\hline
		\textbf{C} & \textbf{FP rate} & \textbf{TP rate} \\
		$10^{-1}$ & 0.00 & 0.0004 \\
		$10^{-0.5}$ & 0.005 & 0.121 \\
		$10^{0}$ & 0.020 & 0.256 \\
		$10^{0.5}$ & 0.026 & 0.309 \\
		$10^{1}$ & 0.033 & 0.0336 \\
		\hline
	\end{tabular}
\end{table}

\section*{F}

We can modify the soft-margin SVM to incorporate different costs for different
forms of misclassification. Our typical set-up assigns the same cost C to margin
violations (e.g., misclassification) for both classes:

$$
\min_{w \in \mathbb{R}^d, w_0 \in \mathbb{R}, \xi \in \mathbb{R}^n} 
\frac{1}{2} ||w||^2 + C \frac{1}{n} \sum_{i=1}^n \xi_i
$$
$$
s.t. \quad Y_i(\langle w, x_i \rangle + w_0) \geq 1 - \xi_i \quad \forall i
$$
$$
\quad \xi_i \geq 0 \quad \forall i
$$

We can instead choose to assign different costs $C_+$ and $C_-$ for margin
violations by the positive and negative classes, respectively.

$$
\min_{w \in \mathbb{R^d}, w_0 \in \mathbb{R}, \xi \in \mathbb{R}^n} 
\frac{1}{2} ||w||^2 + C_+ \frac{1}{n_+} \sum_{i | y_i = 1} \xi_i + 
C_- \frac{1}{n_-} \sum_{i | y_i = -1} \xi_i
$$

In our case, the negative class corresponds to not-spam while the positive class
corresponds to spam. We feel that it is highly undesirable to miss important
email. We can operationalize this by more heavily penalizing margin violations
by the not-spam class.

$$
C_- >> C_+
$$

\begin{thebibliography}{2}
	\bibitem{hsu}
		Chih-Wei Hsu, Chih-Chung Chang, and Chih-Jen Lin.
		\textit{A Practical Guide to Support Vector Classification}.
		2016.

\end{thebibliography}
\end{document}