### 2018-04-02 week7-8/week7.tex
kernel space notes

\documentclass{article}

\begin{document}
A basic way to measure accuracy is the \textbf{confusion matrix}:

\begin{tabular}{r r r r}
$y$ & $\hat{y}$ & & \\
1 & 1 & & True positive \\
1 & -1 & & False negative \\
-1 & 1 & & False positive \\
-1 & -1 & & True negative \\
\end{tabular}

\medskip
Synonyms for TP: \textbf{recall}, \textbf{sensitivity}

\medskip
Synonyms for TN: \textbf{specificity}

\medskip
Positive predictive value = precision = $\frac{TP}{TP+FP}$

\medskip
Negative predictive value = $\frac{TN}{TN+FN}$

\medskip
Misclassification rate = $\frac{FP + FN}{n}$

\medskip
But misclassification rate is bad for unbalanced classes (e.g. not helpful to
always predict that a patient does not have a rare form of cancer). Better
measure: \textbf{F-score}
$$
F = \frac{2 (PPV + TPR)}{PPV + TPR}
$$
\medskip
Also better: ROC curve

\medskip
\underline{Proof of why using validation sets will produce the optimal
model}

\medskip

\textbf{Notation}

\smallskip

\begin{tabular}{l l}
$\{f_\mu\}_{\mu \in \mathcal{M}}$ & Models under consideration \\
$|\mathcal{M}|$ & Cardinality of model class (e.g. how many parameters) \\
$\hat{R}_m()$ & misclassification error of model on validation set of size $m$ \\
$R()$ & expected misclassification error of model \\
$f_{\hat{\mu}}$ & model achieving lowest validation error \\
\end{tabular}

\medskip

\textbf{What we want to show}: with probability of at least $1-\delta$,
$$
R(f_{\hat{\mu}}) \leq \min_{\mu \in \mathcal{M}} R(f_\mu) +
\sqrt{\frac{2}{m}\log\left(\frac{2|\mathcal{M}|}{\delta} \right)}
$$

i.e. as $m$ (the size of validation set) grows, the expected error of the model
achieving the lowest validation error attains that of the theoretically optimum
model.

\medskip

\textbf{How we show it}:

\medskip

Start with the difference in terms of expected risk between the model achieving lowest error on the
validation set (henceforth ``empirically best model'')and the model achieving
lowest expected error (henceforth ``theoretically best model'').

$$
R(f_{\hat{\mu}}) - \min_{\mu \in \mathcal{M}} R(f_\mu)
$$

Add in canceling terms reflecting the empirical error on validation set of size
$m$ of the empirically best model.

$$
= R(f_{\hat{\mu}}) -  \hat{R}_m(f_{\hat{\mu}}) + \hat{R}_m(f_{\hat{\mu}})- \min_{\mu \in \mathcal{M}} R(f_\mu)
$$

Because we have defined $f_{\hat{\mu}}$ to be the empirically best model, the
difference between its empirical risk and the expected risk of the theoretically
best model must be less than or equal to the largest possible empirical vs.
theoretical risk gap (i.e., the model $\mu$ producing the greatest difference
between its empirical and expected risk).

$$
\leq R(f_{\hat{\mu}}) -  \hat{R}_m(f_{\hat{\mu}}) + \max_{\mu \in
\mathcal{M}}(\hat{R}_m(f_\mu) -  R(f_\mu))
$$

And again by definition, the empirical vs. theoretical risk gap of the empirically best model
must be less than the empirical vs. theoretical risk gap of the model with the
largest such gap.

$$
\leq 2 \max_{\mu \in \mathcal{M}}|\hat{R}_m(f_\mu) -  R(f_\mu)|
$$

Now, consider the probability of that largest possible empirical vs. expected error gap 
being greater than some arbitrary amount:

$$
P( \max_{\mu \in \mathcal{M}}|\hat{R}_m(f_\mu) -  R(f_\mu)| > \epsilon)
$$

There may be multiple possible models producing that event:

$$
= P\left(\bigcup_{\mu \in \mathcal{M}} \{ |\hat{R}_m(f_\mu) -  R(f_\mu)| >
\epsilon)\}\right)
$$

Boole's inequality states that $P\left( \cup_i A_i \right) \leq \sum_i
P(A_i)$. Thus, the previous inequality must be less than or equal to its summed
version

$$
\leq \sum_{\mu \in \mathcal{M}} P(|\hat{R}_m(f_\mu) -  R(f_\mu)| > \epsilon)
$$

Hoeffding's inequality is $P(|\bar{S}_m - E(\bar{S}_m| \geq \epsilon) \leq 2
\exp (-2m\epsilon^2 / (b-a)^2)$. We are using Bernoulli variables (error
calculating using indicator function for whether prediction is correct or not)
so $a-b=1$. Thus:

$$
P(|\hat{R}_m(f_\mu) -  R(f_\mu)| > \epsilon) \leq 2 \exp (-2m\epsilon^2)
$$

We plug this back into the previous summation:

$$
\sum_{\mu \in \mathcal{M}} P(|\hat{R}_m(f_\mu) -  R(f_\mu)| > \epsilon) \leq  \sum_{\mu \in \mathcal{M}} 2 \exp (-2m\epsilon^2)
$$
$$
=2|\mathcal{M}|\exp(-2m\epsilon^2)
$$

Use  
$$\epsilon = \sqrt{\frac{\log \left( \frac{2|\mathcal{M}|}{\delta}\right)}{2m}} $$

Then, we plug this value into our previous expression
$$
2|\mathcal{M}|\exp(-2m\epsilon^2)
$$
$$
= 2 |\mathcal{M}|\exp \left( \frac{-2m \log \left( \frac{2|\mathcal{M}|}{\delta}\right)}{2m}
\right)
$$
$$
= \delta
$$

Going all the way back to the start, we now can say

\bigskip

\textbf{Mallow's CP}

Is a thing, but don't feel compelled to take detailed notes on it

\end{document}