### 2018-04-13 hw5/hw5.tex
rover 2.0

\documentclass{article}

\usepackage{bm}

\usepackage{geometry}
\geometry{margin=1.2in}

\title{STAT 672 HW 5}

\author{Tom Wallace}

\begin{document}

\maketitle

\section*{I-D}

10-fold cross validation error is 0.19

This error rate is lower than we would expect based on the data-generating
process described in \textbf{1A}. 
There is no true relationship between $\bm{X}_i$ and $Y_i$ and so in theory the logistic
regression model should be no more effective than random guessing. We thus would
expect a cross-validation error of about 0.5. The better-than-expected performance 
is attributable to two factors.

One, $D >> n$. Simply by random chance some features will be highly correlated with the
label. Our regularization process selected these features to be part of the
model. If $D$ is not so large relative to $n$, model performance drops to the
expected level. For example, I ran the simulation with $n=10000$ and $D=100$,
and obtained a cross-validation error of 0.49.

Two, we violated the central premise of cross-validation---a strict segregation
between training and validation sets---by conducting some model training 
on the full data-set. The regularization process in \textbf{1B} selected a
subset of features $\bm{X_S}$ with a particularly strong relationship with label
$\bm{Y}$; we only conducted cross-validation \textit{after} this feature
selection was performed on the \textit{whole} dataset. In this sense we
``cheated'' and allowed the training process to incorporate information
from the validation set. A correct implementation of cross-validation would
conduct the regularization process of \textbf{1B} only on the training set, and
doing so would cause the better-than-expected performance to disappear.

\section*{II-A}

The best models and the associated cross-validation error is given below. The
numbers in the left-hand column are the indices of features included in the
model.

\begin{table}[h!]
	\centering
	\begin{tabular}{|l r|}
		\hline
		\textbf{Features in Model} & \textbf{5-Fold CV Error} \\
		0  3  4  5  6  9 11 14 & 0.225 \\
		0  3  4  5  6  9 11 15 & 0.225 \\
		1  3  4  5  6  8  9 13 14 & 0.225 \\
		1  3  4  5  6  8  9 13 & 0.225 \\
		1  3  8  9 10 11 13 14 & 0.225 \\
		1  3  8  9 10 11 13 & 0.225 \\
		2  3  9 10 14 & 0.225 \\
		2  3  9 10 & 0.225 \\
		3  4  5  6  9 13 14 & 0.225 \\
		3  4  5  6  9 13 & 0.225 \\
		3  6  7  8  9 10 14 & 0.225 \\
		3  6  7  8  9 10 & 0.225 \\
		\hline
	\end{tabular}
\end{table}

\section*{II-B}

The final classifier\footnote{As shown in the preceding table, there are
multiple models that achieve the lowest cross-validation error, so really this
sentence should be final classifier\textit{s}. I tried several, and all led to
the same phenomenon.} trained on the whole training data attains an error rate of
0.502 against the testing data (\texttt{Xtest, Ytest}). This is much higher
than the cross-validation error of 0.225 obtained in \textbf{II-B}, and in fact
is about what one would get from random guessing. There are several reasons for
this underperformance.

One, $n_{\mathrm{train}} << n_{\mathrm{test}}$. Because the model is trained on
such a small dataset, it did not learn the general relationship between features
and label. Note that the mathematical proof of why validation works (slides 28-29 
of class 6) implies that the probability of producing the optimal model
increases with the size of the validation set. It thus is
unsurprising that our small validation set failed to produce a good model.

Two, our classifier lacks a regularization parameter, and so it overfits to the
training data. This results in achieving low empirical error (i.e. good
performance on the training set) but high generalization error (i.e. poor
performance on the training set).

\section*{II-C}

Randomly permuting the labels in \texttt{Ytrain} could expose the aforementioned
issues without the need for a test set. Randomly permuting the labels would mean
there is no true relationship between feature and labels, and so a model attempting
to predict label based on features should be no better than random guessing
(i.e. should achieve cross-validation error of about 0.5). 
If the model achieves much lower error than this,
there is something wrong with the feature selection and/or classifier.

\section*{II-D}

\end{document}