### 2018-04-15 paper/paper.tex
rover 2.0

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

\usepackage[citestyle=authoryear]{biblatex}
\addbibresource{paper.bib}

\author{Tom Wallace}

\title{STAT 672 Final Project: Stochastic Gradient Descent}

\begin{document}

\maketitle

\section{Introduction}

\subsection{Organization}

This paper is divided into four sections. The remainder of this
\textbf{Introduction} section gives intuitive motivation for stochastic gradient
descent (SGD). The \textbf{Method and Theory} section more rigorously presents the
mathematics of SGD and some of its notable properties. The
\textbf{Applications} sections highlights the real-world settings and uses of SGD,
including a case study data analysis. The \textbf{Conclusion} section summarizes
overall findings.

\subsection{Motivation}

Optimization is fundamental to statistical modeling. The chief task of
statistical modeling is to characterize the relationship between 
explanatory variables and an outcome variable, and the chief method for doing so
is to estimate values for coefficients that best relate
each explanatory variable to the outcome variable. The term ``best''
implies picking coefficient values that
maximize some measure of goodness (e.g. likelihood) or minimize some measure of
badness (e.g. loss function). 
Mathematical optimization is the typical
route to achieving such minimization or maximization. Two important
considerations for optimization are parametric assumptions and computational
complexity. SGD, an optimization technique, is
particularly motivated by these considerations.

\subsubsection{Parametric vs. non-parametric}

Assuming that the outcome variable follows a
particular statistical distribution aids the computation of optimal coefficients. 
For example, assumptions in ordinary least squares (OLS)
regression---assumptions that readers almost certainly are familiar with and so will not be
repeated here---allow a closed form solution. The optimization problem of
choosing coefficient $\hat{\bm{\beta}}$ that minimize squared error is solved
by $\hat{\bm{\beta}} = (\bm{X}'\bm{X})^{-1}\bm{X}'\bm{Y}$ (where $\bm{X}$ is the
feature matrix and $\bm{Y}$ the outcome variable).

Even if a parametric model does not have a closed-form solution, the parametric
assumption allows some useful optimization techniques. Consider logistic
regression. The maximum likelihood estimator (MLE) approach for estimating
coefficients leads to a system of $D$ equations. This system of
equations typically is numerically solved using the iterative Newton-Raphson
algorithm:

$$
\hat{\bm{\beta}}_{n+1} = \hat{\bm{\beta}}_{n} -
\bm{H}^{-1}(\hat{\bm{\beta}}_n)\bm{J}(\hat{\bm{\beta}}_n)
$$

$\bm{J}$ is the Jacobian (the
first derivative of the log-likelihood function $l$ with respect to each $w_j$)
and $\bm{H}$ is the Hessian (the second derivative of $l$ with respect to $w_j,
w_{j'}$). The practicality of Newton-Raphson thus depends on whether it is convenient to
find $\bm{J}$ and $\bm{H}$. It is convenient for logistic regression
because parametric and independent-and-identically-distributed (IID) assumptions
mean $l$ is a simple sum of the log probability distribution
function (PDF, in this case binomial) for each observation. We ``know'' (assume)
the form of this PDF and so are
confident that the second derivative exists and is not too onerous to calculate.
In non-parametric settings, we often cannot be so certain and face the possibility of
$\bm{H}$ being non-existent or cumbersome.

The need to conduct optimization in non-parametric settings is a chief
motivation for gradient descent (GD), of which SGD is a variant. In
non-parametric settings---most notably supervised and unsupervised statistical
learning, in which we again seek to find optimal coefficients to relate input
variables to output variables for the purposes of classification or regression---there 
typically is no closed form solution for the coefficients. It
also may not be convenient to find and evaluate the Hessian, making
Newton-Raphson undesirable. SGD does not require any parametric assumptions. 
In its most basic form, SGD only requires finding the gradient (though some extensions
do need the Hessian or an approximation to it). 
SGD thus is well-suited for non-parametric settings.

\subsubsection{Computational Complexity}

How an optimization technique scales with sample size $n$ is another important
consideration. It is little comfort if a method reaches the correct solution but
requires an excessive amount of time to do so. ``Plain'' or
``batch'' GD requires evaluating the gradient for every single observation,
every single iteration, until the algorithm converges. For example, for a
dataset of $n=10^6$ that required 25 iterations to converge, batch GD would require 
evaluating the gradient $25 \times 10^6$ times. This scaling with
$n$ can cause untenably long computation time.

SGD alleviates these computational difficulties by requiring the gradient to be
evaluated for only a single randomly chosen observation per iteration. This
approach means convergence is ``noisier'' and hence requires more iterations to
converge, but each iteration is less complex to compute and so can be done
faster. SGD thus scales much more favorably with $n$ than GD, and so is
particularly useful for large-$n$ applications such as machine learning
and big data problems.

\section{Method and Theory}

\subsection{Basic setup}

Loosely following the set-up of \cite{bottou2010large}, consider a typical supervised classification problem. We have
feature matrix $\bm{X}_{n \times D, n \in \mathbb{R}^n, D \in \mathbb{R}^D}$ (corresponding to $n$ observations and $D$
features) and labels $\bm{Y}_{n \times 1}, Y_i \in \{-1, 1\}$. The goal is to
predict a particular observation's label $Y_i$ using that observation's features $\bm{X}_i$. 
We have a hypothesis class $\mathcal{F}$ consisting of
various functions $f_{\bm{w}}(\bm{X}_i) \in \mathcal{F}$ parametrized by weight
vector $\bm{w}$. We have loss function $L(Y_i, f_{\bm{w}}(\bm{X}_i))$ that
expresses the cost of mis-classification. Assume for now that $L$ is convex. 
We will consider the optimal function $\hat{f}_{\bm{w}}(\bm{X}_i) \in F$ 
that which minimizes empirical risk over all observations: $\frac{1}{n}
\sum_{i=1}^n L(Y_i, f_{\bm{w}}(\bm{X}_i))$.
Denote $\hat{\bm{w}}$ the weight coefficients of this optimal function. We thus
have:

\begin{equation}
	\hat{\bm{w}} = \argmin_{\bm{w}}\frac{1}{n} \sum_{i=1}^n L(Y_i, f_{\bm{w}}(\bm{X}_i))
\end{equation}

Because $L$ is convex, basic calculus tells us that there is one critical point,
it is the global minimum, and it is located where the gradient of the loss
function with respect to $\bm{w}$, $\nabla_{\bm{w}}$, is zero. GD is an iterative
algorithm to numerically approximate this point:

\begin{equation}
	\bm{w}_{t+1} = \bm{w}_t - \gamma \frac{1}{n}\sum_{i=1}^n
	\nabla_{\bm{w}}L(Y_i, f_{\bm{w}_t}(\bm{X}_i))
\end{equation}

$t$ refers to iterations. $\gamma$ is a parameter controlling the step
size (also called ``learning rate''). Assume for now that $\gamma$ is fixed. 
The algorithm stops when the gradient is
arbitarily close to zero; or, put differently, when the difference between this
iteration's estimate and the last iteration's estimate is arbitrarily small.

\begin{equation}
	\bm{w}_t - \bm{w}_{t-1} \leq \epsilon
\end{equation}

Note that the batch GD algorithm presented in (2) requires evaluating the
gradient for every single observation $i$. In a large-$n$ dataset, this can be
computationally infeasible. SGD's innovation is to instead evaluate only a
single randomly chosen observation $i$ at each iteration $t$, greatly lightening
the computational load.

\begin{equation}
	\bm{w}_{t+1} = \bm{w}_t - \gamma 
	\nabla_{\bm{w}}L(Y_i, f_{\bm{w}_t}(\bm{X}_i))
\end{equation}

Having outlined the basic form of SGD, we now examine some of its key
properties.

\subsection{Key Properties}

\subsubsection{Correctness}

\subsubsection{Speed}

\subsection{Extensions}

The basic SGD algorithm has been extended in different ways. The popularity of
the algorithm disallows a comprehensive or detailed treatment of all
development. This sub-section covers some of the more interesting extensions.

\cite{bottou2012stochastic}
\cite{boyd2004convex}
\cite{dal2015calibrating}
\subsubsection{Step Size (Learning Rate)}

\cite{shalev2011pegasos}

\subsubsection{Momentum}

\cite{polyak1992acceleration}
\cite{nesterov}

\subsubsection{Averaging}

\subsubsection{Predictive Variance Reduction}

\subsubsection{Parallelization}

SGD is commonly used in large-$n$, computationally demanding applications. Thus,
even though SGD is a computational improvement over batch GD, there has been
interest in whether SGD can be made even faster by parallelizing it.
\cite{zinkevich2010parallelized} present novel algorithms for doing so. The
actual algorithms are strikingly simple; their proof is highly technical and
omitted here.

The parallelization technique essentially is averaging. In line with previous
notation, suppose we have fixed learning rate $\gamma$,

Advantages

\section{Applications}

\subsection{SGD and Statistical Learning}

\section{Conclusion}

\printbibliography

\end{document}